Building socket-aware BPF programs

### 2018-11-05 paper/paper.tex
paper: Minor fixups

The BPF verifier must ensure that all BPF programs that are loaded into the
kernel are safe to execute. This necessitates verification of every path
through the program to ensure constraint such as that the program will halt,
and that any dereferencing of pointers is done in a safe and controlled manner.
For accessing socket pointers, this requires validation that memory accesses
are within the bounds of the memory allocated for the structure, and that no
modifications are made to the socket fields that would be cause the core socket
handling logic to get into a bad state. This is provided through extension of
existing functionality which performs bounds verification and offset conversion
for programs operating on CGroups hooks. Additionally, the underlying memory
must remain associated with the socket for the duration of accesses from the
BPF program, which may be guaranteed by taking a reference on sockets for the
duration of their use. The following sections explain how this is implemented
through the introduction of a new pointer type in the verifier, and through the
extension of the verifier to recognise assembly instructions that represent
reference acquisition and release.

@@

Another aspect of extensibility is the ability to lookup sockets which are not
TCP or UDP. The initial RFC of this patch series proposed a single lookup
function which would choose the Layer~4 protocol based upon a field in the
tuple~\cite{sk-lookup-rfc}, however this would make it more difficult for BPF
program writers to detect the support for different protocol types at compile
time. When a socket type is unsupported, the initial implementation would
return \verb+NULL+ at runtime, implying that there is no such socket. There is
not a significant number of supported Layer~4 socket types in the Linux stack
today, and the number is not expected to increase drastically, so it was
considered simpler and easier to split out each Layer~4 socket lookup function
into an independent helper call, so that the result could simply return a
socket or \verb+NULL+ and not need to encode a representation of other error
conditions. This simplifies not only the API, but also the verifier
implementation that must understand the return type of the helper functions.

@@

In cases where Linux is configured as a firewall between applications that
exist in distinct Linux kernel copies running in the network, this is
intuitive. The information that is needed to perform stateful firewalling is
not present on the Linux instance that needs this information, so to understand
the state of the peers the information must be either shared from the peers
(assuming that the firewall orchestrator has control over the peer endpoints),
or this information must be reconstructed based upon the packets that are seen
on this intermediate Linux instance. However, in many cases this model is used
when filtering traffic where one peer is co-located with the Linux stack that
performs the firewalling, which already contains information about the local
end of the connection. In this use case, the connection tracking table
effectively duplicates knowledge that Linux already creates and maintains.

@@

The author would like to acknowledge Alexei Starovoitov, Daniel Borkmann, Jarno
Rajahalme, Martin KaFai Lau, Nitin Hande and Thomas Graf for guidance, review
and testing of the patches, and feedback on this paper.

>>>>

The BPF verifier must ensure that all BPF programs that are loaded into the
kernel are safe to execute. This necessitates verification of every path
through the program to ensure constraint such as that the program will halt,
and that any dereferencing of pointers is done in a safe and controlled manner.
For accessing socket pointers, this requires validation that memory accesses
are within the bounds of the memory allocated for the structure, and that no
modifications are made to the socket fields that would cause the core socket
handling logic to get into a bad state. This is provided through extension of
existing functionality which performs bounds verification and offset conversion
for programs operating on CGroups hooks. Additionally, the underlying memory
must remain associated with the socket for the duration of accesses from the
BPF program, which may be guaranteed by taking a reference on sockets for the
duration of their use. The following sections explain how this is implemented
through the introduction of a new pointer type in the verifier, and through the
extension of the verifier to recognise assembly instructions that represent
reference acquisition and release.

@@

Another aspect of extensibility is the ability to lookup sockets which are not
TCP or UDP. The initial RFC of this patch series proposed a single lookup
function which would choose the Layer~4 protocol based upon a field in the
tuple~\cite{sk-lookup-rfc}, however this would make it more difficult for BPF
program writers to detect the support for different protocol types at compile
time. When a socket type is unsupported, the initial implementation would
return \verb+NULL+ at runtime, implying that there is no such socket. Only a
few Layer~4 socket types exist in the Linux stack today, and the number is not
expected to increase drastically, so it was considered simpler and easier to
split out each Layer~4 socket lookup function into an independent helper call,
so that the result could simply return a socket or \verb+NULL+ and not need to
encode a representation of other error conditions. This simplifies not only the
API, but also the verifier implementation that must understand the return type
of the helper functions.

@@

In cases where the kernel is configured as a firewall between applications that
exist in distinct kernel instances running in the network, this is intuitive.
The information that is needed to perform stateful firewalling is not present
on the kernel instance that needs this information, so to understand the state
of the peers the information must be either shared from the peers (assuming
that the firewall orchestrator has control over the peer endpoints), or this
information must be reconstructed based upon the packets that are seen on this
intermediate kernel instance. However, in many cases this model is used when
filtering traffic where one peer is co-located with the kernel stack that
performs the firewalling, which already contains information about the local
end of the connection. In this use case, the connection tracking table
effectively duplicates knowledge that the kernel already creates and maintains.

@@

The author would like to acknowledge Alexei Starovoitov, Daniel Borkmann, Jarno
Rajahalme, John Fastabend, Martin KaFai Lau, Nitin Hande and Thomas Graf for
guidance, review and testing of the patches, and feedback on this paper.
